  title: "Minor thesis-SPATIO-TEMPORAL ANALYSIS OF SMART METERS DATA-  Solar power forecasting."
output:
  html_document:
  df_print: paged
---

## load libraries
```{r}
library(data.table)
library(zoo)
library(xts)
library(dtw) 
library(forecast) 
library(randomForest)
library(reshape2)
library(dplyr) 
library(Metrics)
```

##Step 1: Read "Customer solar output dataset" and "Postcode cloud cover dataset".

#### 1.1 read customer data (ID, Postcode)
```{r}
c_pos <- read.csv("Postcode.csv")
head(c_pos)
```

#### 1.2 unique Postcodes
```{r}
pos <- sort(unique(c_pos$Postcode))
pos
```

#### 1.3 cloud cover postcodes' idx
```{r}
cc_pos_idx <- 
  "3 5 6 8 13 15 16 17 19 20 21 23 24 25 26 27 30 31 32 37 38 39 40 41 43 44 47 48 49 51 52 53 54 55 56 57 58 60 61 66 67 68 69 70 71 72 73 76 77 78 80 81 82 83 84 85 86 87 88 89 90 91 92 93 94 95 98 99"
cc_pos_idx <- as.numeric(unlist(strsplit(cc_pos_idx, " ")))
cc_pos_idx
```

#### 1.4 cloud cover postcodes
```{r}
cc_pos <- c()
for (i in cc_pos_idx){
  cc_pos <- c(cc_pos, pos[i])
}
cc_pos
```

#### 1.5 cloud cover xts
```{r}
cloud <- read.csv(file = "cloudCover_df.csv", head = T)
cloud_zoo <- read.zoo(cloud ,header = T)
cloud_xts <- as.xts(cloud_zoo)
head(cloud_xts,2)
```

#### 1.6 solar xts 
```{r}
solar <- read.csv(file = "solar_xts_adjusted.csv", head = T)
solar$date<- as.POSIXct(solar$date, format="%Y-%m-%d %H:%M:%S",tz = "UTC")
solar_zoo <- read.zoo(solar ,header = T)
solar_xts <- as.xts(solar_zoo)
head(solar_xts,2)
```

##Step 2. DTW mapping function that generates mapping data as external inputs for predictive models for a given postcode with its neighbouring postcode.

#### 2.1 Function: DTW 
```{r}
dtw_trace <- function(c1,c2,step) {
  trace <-dtw(c1, c2, k=T, step=step) 	
}
```

#### 2.2 Function: DTW Similarity (distance)
```{r}
dtw_calculation <- function (c_pos, nbs_pos, cloud_xts) {
  
  #for storing the DTW distance of each pair
  dist <- as.data.frame(matrix(0, length(nbs_pos),2))
  colnames(dist) <- c("distance", "postcode")
  
  #postcode's cloud cover series
  c1 <- unclass(cloud_xts[,c_pos])[,] 
  
  #distance = Infinity if c1 is all zero (missing value) 
  if (sum(c1)==0){
    dist[,1] <- Inf ; dist[,2] <- 0
    return (dist)
  }
  else{
    for (i in 1:length(nbs_pos)){
      
      #neighbouring postcode's cloud cover series
      c2 <- unclass(cloud_xts[,nbs_pos[i]])[,] 
      
      #distance = Infinity if c2 is all zero (missing value) 
      if (sum(c2)==0){
        dist[i,1] <- Inf ; 
        dist[i,2] <- nbs_pos[i]
      }
      
      #DTW of c1 to c2, store DTW distance of the pair
      else{ 
        trace <- dtw_trace(scale(c1),scale(c2),typeIa)
        dist[i,1] <- as.numeric(trace$distance)
        dist[i,2] <- nbs_pos[i]
      }  
    }
    #ascending order for postcodes by distance
    dist<- dist[order(dist$distance),] 
    return (dist)
  }
}
```

#### 2.3 Function: DTW Daily Mapping:
3 * 48 matrix, where 3 = leg(mapping index),  dummy (1 for no mapping, 0 otherwise), neighbouring postcode's cloud cover mapped from leg

```{r}
dtw_mapping <- function (c_pos, nb_pos, cloud_xts) {
  
  #no mapping found: 
  if (nb_pos == 0){ 
    
    mapping <- matrix(0,nrow(cloud_xts),3)
    mapping[,2] <- 1
    hd <- matrix(0,9,3) ; hd[,2] <- 1
    tl <- matrix(0,8,3) ; tl[,2] <- 1
    
    #combine (t1 to t9 , t10 to t40 , t41 to t48) 
    daily_mapping <- rbind(hd, mapping, tl) 
    return (daily_mapping)
  }
  else{   
    
    #DTW for t10 to t40
    c1 <- unclass(cloud_xts[,c_pos])[,]
    c2 <- unclass(cloud_xts[,nb_pos])[,]
    trace <- dtw_trace(scale(c1),scale(c2),typeIa)
    
    a <- 0 ; 
    time_index <- 0
    mapping <- matrix(0,length(c1),3)
    
    #index.vec
    for ( k in 1:length(trace$index1)){
      if (trace$index1[k]!=a){
        time_index <- time_index +1
        mapping[time_index,1] <- trace$index2[k] - trace$index1[k]
      }
      a = trace$index1[k]
    }
    
    #leg.vec = replace idex as 0 forindex > 0
    mapping[mapping[,1]>0,1] <- 0;  
    
    #dummy.vec
    mapping[mapping[,1]==0,2] <- 1
    
    #cloud.vec: cloud cover data c2 mapped by leg.vec 
    idx <- index(mapping) + mapping[,1]
    mapping[mapping[,2]==0,3] <- c2[idx][mapping[,2]==0]
    
    #t1 to t9 (no cloud cover data) 
    hd <- matrix(0,9,3) ; hd[,2] <- 1
    #t41 to t48 (no cloud cover data) 
    tl <- matrix(0,8,3) ; tl[,2] <- 1
    
    #combine all as a daily mapping: combine (t1 to t9 , t10 to t40 , t41 to t48) 
    daily_mapping <- rbind(hd, mapping, tl)
    return (daily_mapping)
  }
}   
```



##Step 3. Define seven methods including (seasonal naive, ARIMA, ARIMA +2 DTW, ARIMA +4 DTW, Random Forest,Random Forest +2 DTW, Random Forest + 4 DTW).

#### 3.1 Arima model 
```{r}
arimaP <- function (pred, predH, train_obv, ct_solar_ts, frequency, K){
  
  pred_matrix <- matrix(0,pred,predH)
  for (i in 1:pred){
    
    #input solar data 
    st <- i ; nd <- i + train_obv-1 
    inputs <- ts(ct_solar_ts[st:nd], frequency = frequency)
    
    #input fourier series
    fuur <- fourier(inputs, K = K)
    fuurf <- fourier(inputs, K= K, h=predH)
    
    #model fitted
    print (i)
    fit<-auto.arima(inputs, xreg=fuur,
                    max.p = 5, max.q = 5,
                    seasonal = FALSE,stepwise = TRUE)
    
    #forecast
    fc <- forecast(fit, xreg=fuurf, h=predH)
    pred_matrix[i,] <- fc$mean
  }
  return (pred_matrix[,predH])
}
```

#### 3.2 Arima DTW model  ( number of DTW subjected to input data :dtw_fn_lag4)
```{r}
arimaDTW <- function (pred, predH, train_obv, ct_solar_ts, frequency, K, dtw_fn_lag4){
  
  pred_matrix <- matrix(0, pred, predH)
  for (i in 1:pred){
    
    #input solar data 
    st <- i ; nd <- i + train_obv-1 
    inputs <- ts(ct_solar_ts[st:nd], frequency = frequency)
    
    #input fourier series
    fuur <- fourier(inputs, K = K)
    fuurf <- fourier(inputs, K= K, h=predH)
    
    #input DTW mapping 
    dtw_train <- dtw_fn_lag4[st:nd,]
    dtw_test <- dtw_fn_lag4[(nd+1):(nd+1+predH-1),]
    
    #model fitted
    print (i)
    fit<-auto.arima(inputs, xreg=cbind(fuur, dtw_train),
                    max.p = 5, max.q = 5,
                    seasonal = FALSE,stepwise = TRUE)
    
    #forecast
    fc <- forecast(fit, xreg=cbind(fuurf, dtw_test), h=predH)
    pred_matrix[i,] <- fc$mean
  }
  return (pred_matrix[,predH])
}
```

### 3.3 Random Forest model
```{r}
RF <- function (pred, predH, train_obv, ct_solar_lags, ct_solar_ts,
                K, lags){
  
  #fourier series 
  fuur <- fourier(ct_solar_ts, K =  K)
  
  #training set------------------
  
  #solar data 
  inputs_train <- ct_solar_lags[1:(train_obv-5-4+1),1:lags]
  outputs_train <- ct_solar_lags[1:(train_obv-5-4+1),(lags+1):ncol(ct_solar_lags)]
  
  #input (solar data, fourier series)
  inputs_train_matrix <- data.table(inputs_train,
                                    fuur[(5+predH):(5+predH+nrow(outputs_train)-1),])

  #test set ---------------------
  
  #solar data 
  st <- train_obv-5+1; nd <- st + pred -1
  inputs_test <- ct_solar_lags[st:nd,1:lags]
  
  #input (solar data, fourier series)
  inputs_test_matrix <- data.table(inputs_test,
                                   fuur[(train_obv+predH):(train_obv+predH+(pred)-1),])
  
  #train and predict---------------------
  print("RF")
  mod <- randomForest(x= inputs_train_matrix,
                      y= outputs_train[,predH], #y = Y(t4)
                      importance = TRUE)
  return(predict(mod,inputs_test_matrix))
} 
```


### 3.4 Random Forest  DTW model ( number of DTW subjected to input data :dtw_fn_lags)
```{r}
RFDTW <- function (pred, predH, train_obv, ct_solar_lags, ct_solar_ts,
                   K, lags, dtw_fn_lags){
  
  #fourier series 
  fuur <- fourier(ct_solar_ts, K =  K)
  
  #training set------------------
  
  #solar data 
  inputs_train <- ct_solar_lags[1:(train_obv-5-4+1),1:lags]
  outputs_train <- ct_solar_lags[1:(train_obv-5-4+1),(lags+1):ncol(ct_solar_lags)]
  
  #DTW mapping 
  dtw_train <- dtw_fn_lags[1:(train_obv-5-4+1),]
  
  #input (solar data, fourier series,DTW mapping)
  inputs_train_matrix <- data.table(inputs_train,
                                    fuur[(5+predH):(5+predH+nrow(outputs_train)-1),],
                                    dtw_train) 
  #test set ---------------------
  
  #solar data 
  st <- train_obv-5+1; nd <- st + pred -1
  inputs_test <- ct_solar_lags[st:nd,1:lags]
  
  #DTW mapping 
  dtw_test <- dtw_fn_lags[st:nd,]
  
  #input (solar data, fourier series,DTW mapping)
  inputs_test_matrix <- data.table(inputs_test,
                                   fuur[(train_obv+predH):(train_obv+predH+(pred)-1),],
                                   dtw_test)
  
  #train and predict---------------------
  print("RF")
  mod <- randomForest(x= inputs_train_matrix,
                      y= outputs_train[,predH], #y = Y(t4)
                      importance = TRUE)
  return(predict(mod,inputs_test_matrix))
} 
```

#### 3.5 Seasonal Naive method
```{r}
s.naive <- function (ct_solar_ts, frequency, pred, train_obv){
  st <- 4 ; nd <- 4+train_obv-1 

  #input solar data 
  inputs <- ts(ct_solar_ts[st:nd], frequency= frequency)
  
  #forecast
  print ("snaive")
  fc <- snaive(inputs, h=pred)
  return (fc$mean)
} 
```



## Step 4. Run implementation on the randomly select dates and output the predictions for the customer. 

#### 4.1 a date list is set for randomly selected start date then plus next 29 day together as 30 days data as the training set.
```{r}
date_ls <- seq(as.Date("2011-01-01"), as.Date("2011-11-30"), by = "1 days") 
head(date_ls)
```

#### 4.2 assign data for running the implementation.
#### Note: for prediction more customer, place the r command(s) shown in Microsoft word file "R command(s) for postcode with neighbours" section 2 into the below block and run it.
```{r}
#run (postcode, neighbouring postcodes, dates, solar data, cloud cover data)

#run prediction for customers under postcod 2134 
run(2134, c(2200,2209,2208,2207,2203,2047,2135,2137,2190,2162,2048,2204), date_ls, solar_xts, cloud_xts,c_pos)

# #run prediction for customers under postcod 2137 
# run(2137, c(2200,2209,2208,2207,2203,2047,2135,2134,2190,2162,2048,2204,2066), date_ls, solar_xts, cloud_xts,c_pos)

# #run prediction for customers under postcod 2203
# run(2203, c(2208,2207,2216,2018,2044,2204,2048,2047,2135,2134,2137,2190), date_ls, solar_xts, cloud_xts,c_pos)
```

#### 4.3 run function for input data 
```{r}
run <- function (x1, x2, date_ls, solar_xts, cloud_xts,customerPostcode){
  #obtain customer ID under the selected postcode 
  custs <- customerPostcode[customerPostcode$Postcode== x1, 1]
  
  c_pos <- paste("X", x1 ,sep="" )
  nb_pos <- paste("X", x2 ,sep="" )
  
  #call s1function
  s1function(custs, c_pos, nb_pos, date_ls, solar_xts, cloud_xts)
}
```

#### 4.4 s1function "for loop" for calling rund function for each randomly select dates and output prediction result when obtained the forecast from rund function
```{r}
s1function <- function (custs, c_pos, nb_pos, date_ls, solar_xts, cloud_xts){
  
  #test on 5 days
  days <- 5 
  # true y + pred y from 7 methods/models
  mds <- 8  
  
  #------------------------------------
  
  #test for all customers under the selected postcode 
  for (cust in custs){
    cust <- paste("X", cust ,sep="" )
    error_marix_all<- matrix(0,0,mds)
    
    #randomly selected days for training then testing  
    set.seed(123)
    seeds <- sample(1:100,days)
    for (seed in  seeds){
      
      #call rund function 
      error_marix <- rund(seed, date_ls, cust, c_pos, nb_pos, solar_xts, cloud_xts, mds)
      
      #get result
      error_marix_all <- rbind(error_marix_all,error_marix)
    }
    name <- paste0("Pred_", cust, ".csv")
    write.csv(error_marix_all, name, row.names=FALSE)
  }
}
```

#### 4.5 rund function for executing the predefined seven methods/models to do the forecast and record the results 
```{r}
rund <- function (seed, date_ls, cust, c_pos, nb_pos, solar_xts, cloud_xts, mds){
  
  frequency <- 48  # 30 mins interval
  lags <- 5        # lags used 
  K <- 5           # fourier 
  
  train_obv <- 48*30*1 #one month training 
  predH <- 2*2     # predict horizon
  pred <- 48*1    #testing Head 
  
  
  #record true y and prediction from the seven methods/models
  error_marix <- matrix(0, pred, mds)
  
  #period for training and testing (seed: randomly selected date for testing )
  set.seed(seed)
  start <- sample(date_ls, 1) ; end <- start+32 
  date_ls <- seq(start,end, by = "1 days") 
  time_frame <- paste(start,end, sep = "/")
  solar_xts_sub <- solar_xts[time_frame]
  ct_solar <- unclass(solar_xts_sub[,cust])[,]
  ct_solar_ts <- ts(ct_solar, frequency= frequency)
  
  #get 4 DTW mapping --------------------
  dtws <- 4
  dtw_ls <- list()
  
  for (i in 1:length(date_ls)){
    cloud_data <- cloud_xts[toString(date_ls[i])]
    dist <- dtw_calculation(c_pos, nb_pos, cloud_data)
    
    temp <- list()
    for (j in (1:dtws)){ 
      nb_po <- dist[j,2]
      temp[[j]] <- dtw_mapping (c_pos, nb_po, cloud_data)
    }  
    mapping <- do.call(cbind, temp)
    dtw_ls[[i]] <- mapping
  }
  dtw_fn <- do.call(rbind, dtw_ls)
  
  #leg4 of 4 DTW mapping (used in ARIMA)-------------
  lag4 <- matrix(0,4,(3*dtws)) 
  lag4[,2]<- 1; lag4[,5]<- 1; lag4[,8]<- 1; lag4[,11]<- 1
  dtw_fn_lag4 <- rbind(lag4,dtw_fn)
  
  
  #leg1 to leg5 of 4 DTW mapping (used in Random forest)----------------
  getLags <- function (vals, lags){#Function to get lags matrix 
    vals <- ts(vals, frequency = 48)
    lagVals <- list()
    for(x in 1:(lags)) {
      lagVals [[x]] <- stats::lag(vals, x)
    }
    mydata <- do.call(cbind, lagVals)
    colnames(mydata) <- NULL
    mydata <- mydata[complete.cases(mydata),]
  }
  
  temp <- list ()
  i1 <- 1
  for (k in seq(1,10,3)){
    dtw_ <- getLags(dtw_fn[,k],lags)
    dtw_missing <- getLags(dtw_fn[,(k+1)],lags)
    dtw_c2 <- getLags(dtw_fn[,(k+2)],lags)
    temp[[i1]] <- cbind(dtw_,dtw_missing,dtw_c2)
    i1 <- i1+1
  }
  dtw_fn_lags <- do.call(cbind, temp)
  colnames(dtw_fn_lags) <- c(paste0("DTWlg[", seq(1:(lags*3*dtws)), "]")) 
  
  #lag values for solar input data----------------
  lagVals <- list()
  for(x in 1:(lags+predH)) {
    lagVals [[x]] <- stats::lag(ct_solar_ts, x)
  }
  ct_solar_lags <- do.call(cbind, lagVals)
  colnames(ct_solar_lags) <- NULL
  ct_solar_lags <- ct_solar_lags[complete.cases(ct_solar_lags),]
  
  #record true y and pred y -----------------------
  #ture y
  st <- train_obv ; nd <- st + pred -1
  ct_solar_lags_testY <- ct_solar_lags[(st):nd, 2:(2+predH-1)]
  error_marix[,1] <- ct_solar_lags_testY[,4]
  
  #pred y from ARIMA
  error_marix[,2] <- arimaP(pred, predH, train_obv, ct_solar_ts, frequency,
                            K)
  #pred y from ARIMA + 2DTW
  error_marix[,3] <- arimaDTW(pred, predH, train_obv, ct_solar_ts, frequency,
                              K, dtw_fn_lag4[,1:(3*2)])
  
  #pred y from ARIMA + 4DTW
  error_marix[,4] <- arimaDTW(pred, predH, train_obv, ct_solar_ts, frequency,
                              K, dtw_fn_lag4[,1:(3*4)])
  
  #pred y from Random Forest  
  error_marix[,5] <- RF(pred, predH, train_obv, ct_solar_lags, ct_solar_ts,
                        K, lags)

  #pred y from Random Forest + 2DTW 
  error_marix[,6] <- RFDTW(pred, predH, train_obv, ct_solar_lags, ct_solar_ts,
                           K, lags, dtw_fn_lags[,1:(3*2*lags)])
 
  #pred y from Random Forest + 4DTW 
  error_marix[,7] <- RFDTW(pred, predH, train_obv, ct_solar_lags, ct_solar_ts,
                           K, lags, dtw_fn_lags[,1:(3*4*lags)])
  
  #pred y from snaive
  error_marix[,8] <- s.naive(ct_solar_ts, frequency, pred, train_obv)
  
  return (error_marix)
}  
```

## End.

